---
title: "Project"
author: "Neetika Saxena"
date: "9/21/2020"
output:
  pdf_document: default
  html_document: default
---

```{r}
knitr::opts_chunk$set(echo = FALSE)
options(scipen=999)
```
## Avocado Analysis
Source: https://www.kaggle.com/timmate/avocado-prices-2020?select=avocado-updated-2020.csv

Question for Analysis:

* What happened to avocado prices and consumption in 2020?
* How do the prices of avocados vary by region?
# Loading Packages
```{r}
if(!require("ggplot2")) install.packages("ggplot2")
if(!require("knitr")) install.packages("knitr")
if(!require("GGally")) install.packages("GGally")
if(!require("gridExtra")) install.packages("gridExtra")
if(!require("dplyr")) install.packages("dplyr")
if(!require("lubridate")) install.packages("lubridate")     
library(ggplot2)
library(knitr)
library(GGally)
library(gridExtra)
library(dplyr)
library(lubridate)
```
# Loading Data
```{r}
av = read.csv("avocado-updated-2020.csv")
```
# Get Duplicate Rows and missing values in the DataSet
```{r}
sum(duplicated(av))
sum(is.na(av))
```
# Explore Variables of DataSet
```{r}
str(av)
```
# Converting Date into Date-Type, which is Character Type
```{r}
av$date = mdy(av$date)
class(av$date)
```

```{r}
ggplot(av, aes(x=average_price)) +
geom_histogram()
```
# Lets visualize the Average_Price of Avacado per year
```{r}
#compute average price per year
library(dplyr)
avg_avocados = av %>%
  select(average_price,year) %>%
  group_by(year) %>%
  filter(average_price < quantile(average_price, 0.99))

avg_avocado_box = ggplot(avg_avocados, aes(factor(year), average_price)) + geom_boxplot(aes(fill = factor(year)), width = 0.75) + labs(title="Average Price of Avocados Per Year", x="Year", y="Price ($)") + scale_fill_brewer(palette="Greens")
print(avg_avocado_box)
```
# Lets visualize the Average_Price of Avacado on th basis of Geography
```{r}
sorted = av %>% 
  group_by(geography) %>%
  mutate(avgvolume = mean(total_volume))
plot = ggplot(data=sorted, aes(x=reorder(as.factor(geography), -avgvolume), y=average_price)) +
  geom_point(aes(color=month(date), alpha=0.1, size=total_volume)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  guides(size = "none") +
  guides(alpha = "none")
print(plot)
```
```{r}
unique(av$geography)
```
Box Plot and Histrogram Representation of Types of Avocado
```{r}
ap1_org <- ggplot(aes(y = average_price), data = subset(av, type == 'organic')) +
  geom_boxplot(fill = 'pink') + 
  ylab('Average Price') +
  ggtitle('Organic Avocado Boxplot')

ap2_org <- ggplot(aes(x = average_price), data = subset(av, type == 'organic')) +
  geom_histogram(fill = 'pink') +
  scale_y_continuous(breaks = seq(0,1000,100)) + 
  scale_x_continuous(breaks = seq(0,3.25,.5)) +
  ggtitle('Organic Avocado Histogram')

ap1_con <- ggplot(aes(y = average_price), data = subset(av, type == 'conventional')) +
  geom_boxplot(fill = 'orange') + 
  scale_y_continuous(breaks = seq(0,2.5,.25)) +
  ylab('Average Price') +
  ggtitle('Conventional Avocado Boxplot')

ap2_con <- ggplot(aes(x = average_price), data = subset(av, type == 'conventional')) +
  geom_histogram(fill = 'orange') +
  scale_y_continuous(breaks = seq(0,800,100)) +
  scale_x_continuous(breaks = seq(0,2.25,.25)) +
  ggtitle('Conventional Avocado Histogram')
suppressMessages(grid.arrange(ap1_org, ap2_org, ap1_con, ap2_con, ncol = 2)) 
```
From 2015 to 2020 the Trend of differnt PLU sold in US region for Coventional Type
```{r}
p1 <- ggplot(aes(x = date, y = X4046), data = subset(av, geography == 'Total U.S.' & type == 'conventional')) +
  geom_line(stat = 'summary',size = 1) +
  ggtitle('Average Sales of X4046 (small) Conventional Avocados Type Over Time')

p2 <- ggplot(aes(x = date, y = X4225), data = subset(av, geography== 'Total U.S.' & type == 'conventional')) +
  geom_line(stat = 'summary',size = 1) +
  ggtitle('Average Sales of X4225 (medium) Conventional Avocados Type Over Time')   

p3 <- ggplot(aes(x = date, y = X4770), data = subset(av, geography== 'Total U.S.' & type == 'conventional')) +
  geom_line(stat = 'summary',size = 1) +
  ggtitle('Average Sales of X4770 (large) Conventional Avocados Type Over Time')
grid.arrange(p1,p2,p3, ncol = 1)
```
From 2015 to 2020 the Trend of differnt PLU sold in US region for Organic Type
```{r}
p1 <- ggplot(aes(x = date, y = X4046), data = subset(av, geography == 'Total U.S.' & type == 'organic')) +
  geom_line(stat = 'summary',size = 1) +
  ggtitle('Average Sales of X4046 (small) Organic Avocados Type Over Time')

p2 <- ggplot(aes(x = date, y = X4225), data = subset(av, geography== 'Total U.S.' & type == 'organic')) +
  geom_line(stat = 'summary',size = 1) +
  ggtitle('Average Sales of X4225 (medium) Organic Avocados Type Over Time')   

p3 <- ggplot(aes(x = date, y = X4770), data = subset(av, geography== 'Total U.S.' & type == 'organic')) +
  geom_line(stat = 'summary',size = 1) +
  ggtitle('Average Sales of X4770 (large) Organic Avocados Type Over Time')
grid.arrange(p1,p2,p3, ncol = 1)
```
Correlation Matrix, to find relation among the numerical value
```{r}
if(!require("corrplot")) install.packages("corrplot")
library(corrplot)
library(data.table)
av.dt = data.table(av)
M = cor(av.dt[,!c("date","type","geography","year")])
corrplot(M, method="number",bg="gold2")
```
